Contextual Identity: Towards Usable Privacy

### 2013-03-02 paper.tex
mostly rewritten

Barth, Datta, Mitchell, and Nissenbaum provide a formal model for contextual
integrity of user data~\cite{barth}. The contextual integrity model focuses on
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for modeling privacy legislation such as
HIPPA and COPPA. Two limitations of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information.  ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to user effect as the content of the message.

Many authors have discovered attacks to link different contextual identities to
the same person. Narayanan and Shmatikov present a re-identification technique
to merge anonymized social graphs from different networks and prove that
user identifiers from Netflix and IMDB and Flickr and Twitter can be
linked~\cite{narayanan1,narayanan2}. Lindamood et al.~and Mislove et al.~show
how to infer previously undisclosed information from released social networking
data~\cite{lindamood,mislove}.

@@

\section{Contextual identity violations}
We consider it a contextual identity violation when multiple identities are
linked without user intent, or when the user cannot choose which contextual
identity to assert.

@@

\subsection{User interface errors}
It is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently sent compromising pictures of
himself publicly~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{How do users think about identity?}
It is unlikely that the majority of users are consciously aware of having
multiple contextual identities. Imagine two users at the opposite ends of the
spectrum:
\begin{itemize}
\item A user who participates in no social networks, reads paper newspapers,
uses the Internet for well-defined tasks (e.g., coordinating visits with
friends over email, looking up restaurant reviews) but does not otherwise
engage in discretionary browsing.
\item A savvy user who is active on Facebook, Twitter, Instagram and has
different posting levels, topics, and followers on each, maintains multiple
email accounts and pseudonyms, is online multiple hours daily for both work and
personal reasons. 
\end{itemize}

The second user may be much more aware of contextual identities, but even the
first user would benefit from protection against contextual identity
violations. For example, engaging in a long-term task such as house-buying
often requires the construction of a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
such as cookie data.

Users' contextual identities may be task-oriented, service-oriented,
relationship-oriented, or guided by other principles. The long-time existence
of support for multiple accounts in email clients suggests that many users have
multiple email identifiers, which could be considered as a proxy for identity
if tied to service accounts. Similarly, data suggests that a large minority of
Twitter users have multiple Twitter accounts, and the plethora of cross-posting
or syndication software from one social network to another suggests that many
users take advantage of multiple social networks~\cite{twitter,xpost}.  That is
not to say that all users have the same tolerance or even need for multiple
explicit identities: in fact many users may be exclusively task-oriented and
use the Internet with specific tasks in mind rather than for social interaction
or exploratory surfing\footnote{Even these users are subject to building
implicit identities through the use of tracking techniques, however.}.

\subsection{How do users manage identity?}
\paragraph{Private browsing mode}
Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, the threat
model for using private browsing mode may be other household members, or it may
be a service to which the user wants to limit cookie leakage.

\paragraph{Profile management}
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so difficult
to use that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.} Tang et
al.~show that most people do not take the time to create labels and maintain
friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

\subsection{How can we improve identity management tools?}
If users' contextual identities are primarily task-oriented, it may be
possible to infer tasks from the user's activity. For example, the
\texttt{about:profile} Firefox extension uses the Open Directory Project and
Alexa to categorize URLs that the user has visited~\cite{aboutprofile}.

@@

\paragraph{Mitigating explicit unwanted linkage}
Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}.
Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\paragraph{Controlling Information Flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.
This type of tool (cookie blocker) helps users assert control over audience.

Protocols like Do Not Track are also a good first step to solving this
problem~\cite{dnt}. Although Do Not Track is unenforceable from the client
side, it provides a clear signal to service providers and ad networks that the
user does not want to be tracked. As adoption increases, ignoring this signal
will lead to loss of reputation for service providers that explicitly act
against their users' wishes, helping users assert control over the effects of
their actions.

\paragraph{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in an
ideal position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, a frequent cause
of regret~\cite{wang}.  Although this approach does not help when the service
provider decides to change the audience on the user's behalf, auditing gives
the user a good opportunity to review posts that might cause regrets. In the
case of high-volume users, such a tool could incorporate sentiment analysis to
aid the user in finding particularly problematic content.

Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via tracking techniques.
Collusion is a tool for visualizing cookies~\cite{collusion}, in particular
third-party cookies set by tracking sites which are typically ad networks.
Once the user knows which sites are tracking them, they can choose to block
cookies from those sites.

\paragraph{Remembering to Forget}
Humans don't have perfect memories, and perhaps social networks should not
either~\cite{delete}. Many users already engage in manual auditing and deletion
of old posts~\cite{fbtips2}. A better solution would be to build tools that let
the user manage this more easily through an API: both Twitter and Facebook
provide APIs for post deletion\footnote{Google Plus
offers a read-only API.}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

\paragraph{Visual reminders of context}
To mitigate privacy violations caused by poor user interfaces, the onus must be
on software developers to provide sufficient visual cues of context. For
example, many but not all Google products support multiple accounts. However,
the visual cue to indicate which account is active is a small, easy to ignore
text name at the top right of the page.

Much work has been done for visual verification of domain authenticity in the
context of phishing, but not much has been done for visual verification of
which user credentials are active~\cite{skins}.

@@

\begin{comment}
\section{Acknowledgements}
The authors thank
Lucas Adamski
Ben Adida
Mike Connor
Ed Lee
\end{comment}

%Convert to IEEE?
\bibliographystyle{plain}
\bibliography{paper}
\end{document}

>>>>

Barth, Datta, Mitchell, and Nissenbaum provide a formal model for contextual
integrity of user data~\cite{barth}. The contextual integrity model focuses on
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for modeling privacy legislation such as
HIPPA and COPPA. Two limitations of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information. For example, ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to user effect as the contents.

Many authors have discovered methods to link different contextual identities to
the same person. Narayanan and Shmatikov present a re-identification technique
to merge anonymized social graphs from different networks and prove that
user identifiers from Netflix and IMDB and Flickr and Twitter can be
linked~\cite{narayanan1,narayanan2}. Lindamood et al.~and Mislove et al.~show
how to infer previously undisclosed information from public social networking
data~\cite{lindamood,mislove}.

@@

\section{Contextual identity violations}
In this section we discuss three types of \textit{contextual identity
violations}. We consider a contextual identity violation to be when multiple
identities are linked without user intent, or when the user cannot choose which
contextual identity to assert in a given situation. In these examples the user
may not be conscious of what a contextual identity is: nevertheless each
example illustrates a loss of user control over what aspects of self to
reveal in a given context.

@@

\subsection{User interface errors}
It is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently published compromising pictures
of himself~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{How do users think about identity?}
We hypothesize many users are not consciously aware of having multiple
contextual identities. In order to develop tools to help users, we need to
understand users' mental models of identity and how information is shared on
the Internet. For computer scientists, authentication is tied to identity, and
even in the absence of authentication, using the same device over time
implicitly creates an identity through tracking techniques and local
information. However, these points are far from obvious to a typical user,
especially for users who don't distinguish the web application from the
browser, from the operating system, or from the device.

The computer science community must free itself from software-specific
notions of identity if we are to help users to whom the very concept of
identity is an enigma. To start understanding our users' mental models,
one might start with the questions:

\begin{itemize}
\item How is identity represented online?
\item Is your identity tied to the device you are using?
\item If you check your mail, read news, log on to a social network at a
computer at the public library, does that affect your identity?
\item What does it mean to share devices? If you lend your tablet to your
sister, is she representing herself, or you?
\item If you register for an account on a service, do you expect that to have
an influence on your online identity? What if you never use that account again?
\item If you visit a website, do you expect that to have an influence on your
browsing experience or online identity? If so, for how long?
\item Which parts of your online identity do you expect to be visible to
your housemates, friends, relatives, or employers?
\item Which parts of your identity do you want to share with whom?
\item Which parts of your identity do you want to keep secret from whom?
\item How does your identity change over time?
\end{itemize}

danah boyd has done excellent work surveying youth and social networks, and Pew
Internet is a good source for self-reported phone surveys on privacy and social
media, but there is otherwise a dearth of research in these
areas~\cite{boyd,pew1,pew2,pew3}.

\subsection{How do users manage identity?}

Users who are aware of having multiple identities engage in the following
identity management techniques.

\paragraph{Multiple accounts}
The long-time existence of support for multiple accounts in email clients
suggests that many users have multiple email identifiers, which could be
considered as a proxy for identity if tied to service accounts. Similarly, data
suggests that a large minority of Twitter users have multiple Twitter accounts,
and the plethora of cross-posting or syndication software from one social
network to another suggests that many users take advantage of multiple social
networks~\cite{twitter}.
\begin{comment}
That is not to say that all users have the same
tolerance or even need for multiple explicit identities: in fact many users may
be exclusively task-oriented and use the Internet with specific tasks in mind
rather than for social interaction or exploratory surfing\footnote{Even these
users are subject to building implicit identities through the use of tracking
techniques, however.}.
\end{comment}

\paragraph{Auditing}
Wang et.~al suggest that common techniques for handling regret online include
manual deletion of regrettable posts, self-censoring or delaying
posts that they predict might bring regret~\cite{wang}.

\paragraph{Private browsing mode}
Private browsing mode exists in all major browsers, but are not a complete
solution~\cite{ABBJ10}. There is no standard behavior for private browsing
mode, and interaction with extensions, treatment of cookies, history, and
bookmarks upon entering and exiting this mode are different and serve different
use cases. For example, entering private browsing in Firefox does not clear
cookies, history, or other session data; session data is cleared on exiting
private browsing mode. In contrast, incognito mode in Chrome makes previous
session state unavailable and clears new session state on exit; this mode
may be preferable as a ``Guest mode'' where the device owner does not want the
guest to either view or pollute existing state.

\paragraph{Profile management}
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so unusable
that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.}

\paragraph{Cookie blockers}
Disconnect and ShareMeNot are browser extensions to disallow interacting with
service providers like Google and Facebook unless users explicitly choose to
interact with that site~\cite{disconnect,franzi}.

\subsection{How can we improve identity management tools?}
\paragraph{Task inference}
If users' contextual identities are primarily task-oriented, it may be
possible to infer tasks from the user's activity. For example, the
\texttt{about:profile} Firefox extension uses the Open Directory Project and
Alexa to categorize URLs that the user has visited~\cite{aboutprofile}.

\paragraph{Relationship inference}
Tang et al.~show that most people do not take the time to create labels and
maintain friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

@@

\paragraph{Mitigating explicit linkage}
Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}.
Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\paragraph{Controlling Information Flow}
Do Not Track is an HTTP request header that signals to the website that the
user does not want to be tracked~\cite{dnt}. Although Do Not Track is
unenforceable from the client side, service providers who ignore signal will
lose reputation by acting against their users' explicit wishes.

\paragraph{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience. A browser is in an ideal position to intermediate social network
posts, aggregate them locally, and present them to the user when they desire to
audit their digital footprint.  For example, a user could see all of the
comments and posts they made in the last week and redact content that is
sensitive or too negative, a frequent cause of regret~\cite{wang}.  Although
this approach does not help when the service provider decides to change the
audience on the user's behalf, auditing gives the user a good opportunity to
review posts that might cause regrets. In the case of high-volume users, such a
tool could incorporate sentiment analysis to aid the user in finding
particularly problematic content.

\paragraph{Implicit identity generation}
Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via tracking techniques.  Collusion is a tool for
visualizing cookies~\cite{collusion}, in particular third-party cookies set by
tracking sites which are typically ad networks.  Once the user knows which
sites are tracking them, they can choose to block cookies from those sites.

\paragraph{Visual reminders of context}
To mitigate privacy violations caused by poor user interfaces, the software
must provide sufficient visual cues of context. For example, many but not all
Google products support multiple accounts. However, the visual cue to indicate
which account is active is a small, easy to ignore text name at the top right
of the page.

\paragraph{Expiring posts}
Humans don't have perfect memories, and perhaps social networks should not
either~\cite{delete}. Many users already engage in manual auditing and deletion
of old posts~\cite{fbtips2}. A better solution would be to build tools that let
the user manage this more easily through an API: both Twitter and Facebook
provide APIs for post deletion\footnote{Google Plus
offers a read-only API.}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

\paragraph{Expiring contextual identities}
Engaging in a long-term task such as house-buying
often requires the construction of a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
such as cookie data.

@@

\begin{comment}
\section{Acknowledgements}
The authors thank
Lucas Adamski
Ben Adida
Mike Connor
Chris Karlof
\end{comment}

\bibliographystyle{plain}
\bibliography{paper}
\end{document}
